---
title: "George_Kamundia_Assignment"
author: "Kamundia"
date: "2024-10-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. DATA WRANGLIING CHALLENGE
```{r challenge1.0}
setwd("C:/Users/Anarchy/Documents/Data_Science/CEMA/L4H_sample_data-main")
#install.packages(c("dplyr", "janitor", "lubridate"))
#install.packages("readr")

library(readr)
library(dplyr)
library(janitor)
library(lubridate)

list.files("sample_data")

# Import the data using the exact file names
baseline_household <- read_csv("sample_data/L4H_household_baseline_sample.csv")
baseline_individual <- read_csv("sample_data/L4H_individual_baseline_sample.csv")
baseline_mother <- read_csv("sample_data/L4H_mother_baseline_sample.csv")

baseline_household <- clean_names(baseline_household)
baseline_individual <- clean_names(baseline_individual)
baseline_mother <- clean_names(baseline_mother)

# Filter rows where hh_eligible is 1
baseline_household_filtered <- baseline_household %>% 
  filter(hh_eligible == 1)

#Merging baseline_individual data and baseline mother data 
merged_data <- baseline_individual %>% 
  inner_join(baseline_mother, by = c("number" = "number_0"))

# Final merge
final_data <- merged_data %>% 
  inner_join(baseline_household_filtered, by = "household_id")

#View(final_data)

str(final_data)

# Check the column names in final_data
#colnames(final_data)

# Select the columns of interest
columns_of_interest <- c("reason_for_ineligibility", "rspntgndr", "h_hfrml_eductn", "rspndtmarital", "rspndt_eductn", "maincme")

# Display the first 10 entries for these columns
head(final_data[columns_of_interest], 20)

unique(final_data$reason_for_ineligibility)
unique(final_data$rspntgndr)
unique(final_data$h_hfrml_eductn)
unique(final_data$rspndtmarital)
unique(final_data$rspndt_eductn)
unique(final_data$maincme)


# Recode the variables using the mappings from the data dictionary
final_data <- final_data %>%
  mutate(
    reason_for_ineligibility = case_when(
      reason_for_ineligibility == "1" ~ "No adult occupier >16 years",
      reason_for_ineligibility == "2" ~ "Withdrawal",
      reason_for_ineligibility == "3" ~ "Other reason",
      TRUE ~ "---"  # Default for unmatched cases
    ),
    
    rspntgndr = case_when(
      rspntgndr == "1" ~ "Male",
      rspntgndr == "2" ~ "Female",
      TRUE ~ "---"
    ),
    
    h_hfrml_eductn = case_when(
      h_hfrml_eductn == "1" ~ "No formal education",
      h_hfrml_eductn == "2" ~ "Primary School",
      h_hfrml_eductn == "3" ~ "Secondary school",
      h_hfrml_eductn == "4" ~ "College-graduate",
      h_hfrml_eductn == "5" ~ "Madrassa",
      h_hfrml_eductn == "6" ~ "Other",
      TRUE ~ "---"
    ),
    
    rspndtmarital = case_when(
      rspndtmarital == "1" ~ "Single",
      rspndtmarital == "2" ~ "Married monogamous",
      rspndtmarital == "3" ~ "Married polygamous",
      rspndtmarital == "4" ~ "Divorced/separated",
      rspndtmarital == "5" ~ "Widow(er)",
      TRUE ~ "---"
    ),
    
    rspndt_eductn = case_when(
      rspndt_eductn == "1" ~ "No formal education",
      rspndt_eductn == "2" ~ "Primary School",
      rspndt_eductn == "3" ~ "Secondary school",
      rspndt_eductn == "4" ~ "College-graduate",
      rspndt_eductn == "5" ~ "Madrassa",
      rspndt_eductn == "6" ~ "Other",
      TRUE ~ "---"
    ),
    
    maincme = case_when(
      maincme == "1" ~ "Sale of livestock & livestock products",
      maincme == "2" ~ "Sale of crops",
      maincme == "3" ~ "Trading/business",
      maincme == "4" ~ "Employment (salaried income)",
      maincme == "5" ~ "Sale of personal assets",
      maincme == "6" ~ "Remittance",
      maincme == "7" ~ "Other",
      TRUE ~ "---"
    )
  )

# Check unique values for verification
unique(final_data$reason_for_ineligibility)
unique(final_data$rspntgndr)
unique(final_data$h_hfrml_eductn)
unique(final_data$rspndtmarital)
unique(final_data$rspndt_eductn)
unique(final_data$maincme)

unique(final_data$lvstckown)

#install.packages("tidyr")
library(tidyr)


# Separate the lvstckown column into individual species columns
final_data <- final_data %>%
  separate(lvstckown, into = paste0("animal_", 1:15), sep = " ", fill = "right", remove = TRUE)

unique(final_data$herdynamics)

# Separate the herdynamics column into individual response columns
final_data <- final_data %>%
  separate(herdynamics, into = paste0("herdynamics_", 1:7), sep = " ", fill = "right", remove = TRUE)


# Display the first few rows of the updated dataset
head(final_data)

# Check the column names in final_data
#colnames(final_data)


# Create the new study_arm column based on village names
final_data <- final_data %>%
  mutate(study_arm = case_when(
    village.x %in% c("Lependera", "Gobb Arbelle", "Nahgan-ngusa", "Sulate", 
                     "Saale-Sambakah", "Namarei", "Manyatta Lengima", "Lokoshula", 
                     "TubchaDakhane", "Rengumo-Gargule") ~ "Study arm 1",
    village.x %in% c("Galthelian-Torrder", "Uyam village", "Galthelan Elemo", 
                     "Nebey", "Rongumo_kurkum", "Urawen_Kurkum", "Eisimatacho", 
                     "Manyatta K.A.G", "Ltepes Ooodo", "Lorokushu", "Marti", 
                     "Manyatta Juu West/East", "Lbaarok1") ~ "Study arm 2",
    TRUE ~ "Study arm 3"  # All other villages
  ))

# Check the column names in final_data
#colnames(final_data)

unique(final_data$study_arm)

# Create herd_dynamics object with specified columns
herd_dynamics <- final_data %>%
  select(interview_date = interview_date.x, 
         household_id, 
         study_arm, 
         cwsbrth, 
         shpbrth, 
         goatsbrth, 
         cmlsbrth, 
         calves_death, 
         bulls_death, 
         cows_death, 
         sheep_death, 
         msheep_death, 
         fsheep_death, 
         goats_death, 
         mgoats_death, 
         fgoats_death, 
         camels_death, 
         mcamels_death, 
         fcamels_death, 
         cowsgft, 
         sheepgfts, 
         goatsgft, 
         cmlsgft,
         cowsgvnout,       
         sheepgvnout,       
         goatsgvnout,       
         cmlsgvnout)

# Create monthyear column
herd_dynamics <- herd_dynamics %>%
  mutate(monthyear = format(ymd(interview_date), "%Y-%m"))  # Convert interview_date to Date format

# Replace "---" with NA and convert to numeric for specific columns
herd_dynamics <- herd_dynamics %>%
  mutate(across(c(cwsbrth, shpbrth, goatsbrth, cmlsbrth, 
                  calves_death, bulls_death, cows_death, 
                  sheep_death, msheep_death, fsheep_death, 
                  goats_death, mgoats_death, fgoats_death, 
                  camels_death, mcamels_death, fcamels_death, 
                  cowsgft, sheepgfts, goatsgft, cmlsgft,cowsgvnout,sheepgvnout,goatsgvnout, cmlsgvnout  ), 
                ~ as.numeric(ifelse(. == "---", NA, .))))

# Check the structure to confirm the changes
str(herd_dynamics)

# Calculate the number of animals born, died, gifted, and given out
animal_summary <- herd_dynamics %>%
  group_by(study_arm, monthyear) %>%
  summarize(
    cows_born = sum(cwsbrth, na.rm = TRUE),
    sheep_born = sum(shpbrth, na.rm = TRUE),
    goats_born = sum(goatsbrth, na.rm = TRUE),
    camels_born = sum(cmlsbrth, na.rm = TRUE),
    cows_died = sum(cows_death, na.rm = TRUE),
    sheep_died = sum(sheep_death,msheep_death, fsheep_death, na.rm = TRUE),
    goats_died = sum(goats_death, mgoats_death,fgoats_death, na.rm = TRUE),
    camels_died = sum(camels_death, mcamels_death, fcamels_death, na.rm = TRUE),
    cows_gifted = sum(cowsgft, na.rm = TRUE),
    sheep_gifted = sum(sheepgfts, na.rm = TRUE),
    goats_gifted = sum(goatsgft, na.rm = TRUE),
    camels_gifted = sum(cmlsgft, na.rm = TRUE),
    cows_given = sum(cowsgvnout, na.rm = TRUE),     
    sheep_given = sum(sheepgvnout, na.rm = TRUE),   
    goats_given = sum(goatsgvnout, na.rm = TRUE),
    camels_given = sum(cmlsgvnout, na.rm = TRUE), 
    .groups = 'drop'  # Ungroup after summarizing
  )

print(animal_summary, n = Inf)

# Create a subset of the dataset with the specified variables
subset_herd_dynamics <- animal_summary %>%
  select(
    study_arm, 
    monthyear, 
    cows_born, 
    sheep_born, 
    goats_born, 
    camels_born, 
    cows_died, 
    sheep_died, 
    goats_died, 
    camels_died, 
    cows_gifted, 
    sheep_gifted, 
    goats_gifted, 
    camels_gifted, 
    cows_given, 
    sheep_given, 
    goats_given, 
    camels_given
  ) %>%
  distinct()  # Remove duplicates

# Check the structure of the new subset
str(subset_herd_dynamics)

summary(subset_herd_dynamics)

#View(subset_herd_dynamics)


# Print the first few rows of the subset to verify
head(subset_herd_dynamics)


library(ggplot2)

# Reshape the data to long format
long_data <- subset_herd_dynamics %>%
  pivot_longer(
    cols = c(cows_born, cows_died, cows_gifted, cows_given,
             sheep_born, sheep_died, sheep_gifted, sheep_given,
             goats_born, goats_died, goats_gifted, goats_given,
             camels_born, camels_died, camels_gifted, camels_given),
    names_to = c("species", "event"),
    names_pattern = "(.*)_(.*)",
    values_to = "count"
  ) %>%
  mutate(event = case_when(
    event == 'born' ~ 'Births',
    event == 'died' ~ 'Deaths',
    event == 'gifted' ~ 'Gifts In',
    event == 'given' ~ 'Gifts Out'
  )) %>%
  drop_na(count)  # Remove NA counts for cleaner plotting

# Create the plot
ggplot(long_data, aes(x = monthyear, y = count, fill = species)) +
  geom_col(position = "dodge") +  # Use dodge for side-by-side bars
  facet_grid(study_arm ~ event, scales = "free_y") +  # Create separate panels for each study arm and event
  labs(title = "Frequencies and Changes Over Time in Animal Events",
       x = "Month-Year",
       y = "Count") +
  scale_fill_manual(values = c("cows" = "blue", 
                               "sheep" = "green", 
                               "goats" = "purple", 
                               "camels" = "orange")) +  # Custom colors for species
  theme_minimal() +  # Clean theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability


```


# 2. DATA VISUALIZATION CHALLENGE
```{r DataVisualization}

# Load the libraries
library(dplyr)
library(ggplot2)
library(sf)
library(tmap)
library(viridis)

# Load the teenage pregnancy data
teen_pregnancy_data <- read.csv("https://raw.githubusercontent.com/cema-uonbi/L4H_sample_data/main/table6_teenpregnancybycounty.csv")

# Read the shapefiles into R
shapefile_path <- "C:/Users/Anarchy/Documents/Data_Science/CEMA/L4H_sample_data-main/shapefiles/County.shp"
kenya_map <- st_read(shapefile_path) 

# Check the structure of both datasets to ensure county names match
str(teen_pregnancy_data)
str(kenya_map)

# Merge the datasets by county
kenya_merged <- kenya_map %>%
  left_join(teen_pregnancy_data, by = c("Name" = "County"))  # Adjusting to match column names
# Check the structure of the merged dataset
str(kenya_merged)

# Check for any NA values in the merged dataset
summary(kenya_merged)

# Handle missing values in the Ever_pregnant column
kenya_merged <- kenya_merged %>%
  mutate(Ever_pregnant = ifelse(is.na(Ever_pregnant), 0, Ever_pregnant))

# Check for any NA values in the merged dataset
summary(kenya_merged)

# Create a map showing the percentage of teenagers who have ever been pregnant by county
ggplot(data = kenya_merged) +
  geom_sf(aes(fill = Ever_pregnant), color = "white") +  # Fill by Ever_pregnant
  scale_fill_viridis(option = "plasma", name = "Percentage of Teenagers Ever Pregnant") +  # Customize color scale
  labs(title = "Percentage of Teenagers Who Have Ever Been Pregnant by County in Kenya (2022)",
       subtitle = "Data from the Kenya Demographic Health Survey",
       caption = "Source: Kenya Demographic Health Survey") +
  theme_minimal() +
  theme(legend.position = "bottom")

```
# 3.0 DATA ANALYTICS CHALLENGE
## Logitstic Regression
```{r LogisticRegression}
#Loading necessary libraries
library(tidyverse)
library(lme4)
library(broom)
library(dplyr)
library(readr)
library(car)
library(ggplot2)
library(caret)
library(vip)
library(pROC)
library(naniar)


# Step 1: Load the dataset
#data_url <- "https://raw.githubusercontent.com/cema-uonbi/L4H_sample_data/main/ideal3a.csv"
#ideal_data <- read_csv(data_url)
ideal_data <- read.csv("C:/Users/Anarchy/Documents/Data_Science/CEMA/L4H_sample_data-main/ideal3a.csv")

summary(ideal_data)
#view(ideal_data)
# This plot shows which columns have missing values together
gg_miss_upset(ideal_data)


# Step 2: Data Preparation
# Set ReasonsLoss to 0 where ReasonLoss1 indicates survival
ideal_data$ReasonsLoss[ideal_data$ReasonsLoss1 == "survived"] <- 0
# Convert 'ReasonsLoss1' to binary (1 = died, 0 = survived)
ideal_data$ReasonsLoss1 <- ifelse(ideal_data$ReasonsLoss1 == "died", 1, 0)

# Calculate the mean of ManualPCV for the specific CalfID
mean_manualPCV <- mean(ideal_data$ManualPCV[ideal_data$CalfID == "CA031210334"], na.rm = TRUE)
ideal_data$ManualPCV[ideal_data$CalfID == "CA031210334" & is.na(ideal_data$ManualPCV)] <- mean_manualPCV



# Fill missing Weight based on CalfID
#ideal_data <- ideal_data %>%
#  group_by(CalfID) %>%
#  mutate(Weight = ifelse(is.na(Weight), mean(Weight, na.rm = TRUE), Weight)) %>%
#  ungroup()

# Fill missing Q.Strongyle.eggs using last observation carried forward (LOCF)
#ideal_data <- ideal_data %>%
#  arrange(VisitDate) %>%
#  mutate(Q.Strongyle.eggs = na.locf(Q.Strongyle.eggs, na.rm = FALSE))

# Fill missing Weight based on CalfID, sublocation, and VisitDate NOTE: Acc= 0.25, AUC= 0.75
ideal_data <- ideal_data %>%
  group_by(CalfID) %>%
  mutate(Q.Strongyle.eggs = ifelse(is.na(Q.Strongyle.eggs), mean(Q.Strongyle.eggs, na.rm = TRUE), Q.Strongyle.eggs)) %>%
  ungroup()

summary(ideal_data)

# Convert categorical variables to factors
ideal_data <- ideal_data %>%
  mutate(
    CalfSex = as.factor(CalfSex),
    Education = as.factor(Education),
    Distance_water = as.factor(Distance_water),
    ELISA_mutans = as.factor(ELISA_mutans),
    ELISA_parva = as.factor(ELISA_parva),
    Theileria.spp. = as.factor(Theileria.spp.),
    ReasonsLoss = as.factor(ReasonsLoss),
    sublocation = as.factor(sublocation)
  )



# This keeps only the last record for each CalfID based on VisitDate
latest_data <- ideal_data %>%
  group_by(CalfID) %>%
  filter(VisitDate == max(VisitDate)) %>%
  ungroup()

summary(latest_data)

# Here, we simply drop rows with NA values for simplicity
latest_data <- na.omit(latest_data)

summary(latest_data)

num_vars <- c("Weight", "Age", "ManualPCV", "RecruitWeight", "Q.Strongyle.eggs")


latest_data[num_vars] <- scale(latest_data[num_vars])

#view(latest_data)

# Step 5: Check distribution of the target variable (survived vs died) with counts
ggplot(latest_data, aes(x = as.factor(ReasonsLoss1))) + 
  geom_bar() + 
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +  # Add count labels
  labs(x = "Reasons Loss (0 = Survived, 1 = Died)", y = "Count") +
  theme_minimal()

# Step 6: Check multicollinearity using VIF
vif_model <- glm(ReasonsLoss1 ~ CalfSex + Education + Distance_water + RecruitWeight + 
                   Age + ManualPCV + Theileria.spp. + ELISA_mutans + ELISA_parva + 
                   Q.Strongyle.eggs, 
                 family = binomial(link = "logit"), 
                 data = latest_data)

vif_values <- vif(vif_model)
print(vif_values)

# Step 7: Split the data into training and testing sets (70% training, 30% testing)
set.seed(123)  # For reproducibility
train_index <- createDataPartition(latest_data$ReasonsLoss1, p = 0.7, list = FALSE)
train_data <- latest_data[train_index, ]
test_data <- latest_data[-train_index, ]

# Identify any factors with a single level in train_data
single_level_factors <- sapply(train_data, function(x) is.factor(x) && length(unique(x)) < 2)
print(single_level_factors)


# Step 8: Fit a standard logistic regression model
logistic_model <- glm(ReasonsLoss1 ~ CalfSex + Education +  RecruitWeight + Weight + 
                        Age + ManualPCV + Theileria.spp. + ELISA_parva + 
                        Q.Strongyle.eggs, 
                      family = binomial(link = "logit"), 
                      data = train_data)

# Step 9: Model summary
summary(logistic_model)

# Step 10: Make predictions on the test set
test_data$predicted_prob <- predict(logistic_model, newdata = test_data, type = "response")
test_data$predicted_class <- ifelse(test_data$predicted_prob > 0.5, 1, 0)  # Use 1 for "died"

# Step 11: Confusion matrix
conf_matrix <- confusionMatrix(as.factor(test_data$predicted_class), as.factor(test_data$ReasonsLoss1))
print(conf_matrix)


# Extracting values from confusion matrix
TP <- conf_matrix$table[2, 2]  # True Positives
TN <- conf_matrix$table[1, 1]  # True Negatives
FP <- conf_matrix$table[1, 2]  # False Positives
FN <- conf_matrix$table[2, 1]  # False Negatives

# Calculate precision, recall, and F1 score
precision <- TP / (TP + FP)      # Positive Predictive Value
recall <- TP / (TP + FN)          # Sensitivity
F1_score <- 2 * (precision * recall) / (precision + recall)

# Output the F1 score
cat("F1 Score:", F1_score, "\n")

# Variable Importance
importance <- varImp(logistic_model, scale = FALSE)
print(importance, row.names = TRUE)

# Visualization of variable importance
vip(logistic_model)

# Step 1: Create the ROC curve
roc_curve <- roc(test_data$ReasonsLoss1, test_data$predicted_prob)

# Step 2: Plot the ROC curve# Optionally: Add AUC to the plot
plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)

# Step 3: Calculate and print the AUC
auc_value <- auc(roc_curve)
cat("AUC:", auc_value, "\n")

```

### CalfID as random effect to account for individual variations among calves
mixed_model <- glmer(ReasonsLoss1 ~ CalfSex + Education + Distance_water + RecruitWeight + 
                       Age + ManualPCV + Theileria.spp. + ELISA_mutans + ELISA_parva + 
                       Q.Strongyle.eggs + (1 | CalfID), 
                     family = binomial(link = "logit"), 
                     data = train_data)

this code showed that:
Given the zero variance for CalfID, it seems a mixed-effects model may not be effective here, as the individual calf differences are not meaningful for this outcome. This suggests that previous observations for a calf are not contributing distinct information toward predicting the final outcome in a way that the random effect can capture.
This influenced the approach of using a simpler logistic regression by only looking at the last VisitDate.



## LinearRegression

```{r LinearRegression}
# Load necessary libraries
library(lme4)
library(dplyr)
library(ggplot2)
library(car)
library(broom.mixed)
library(MuMIn)


# Step 1: Load the dataset
ideal_data <- read.csv("C:/Users/Anarchy/Documents/Data_Science/CEMA/L4H_sample_data-main/ideal3a.csv")

# Set ReasonsLoss to 0 where ReasonLoss1 indicates survival
ideal_data$ReasonsLoss[ideal_data$ReasonsLoss1 == "survived"] <- 0

# Convert 'ReasonsLoss1' to binary (1 = died, 0 = survived)
ideal_data$ReasonsLoss1 <- ifelse(ideal_data$ReasonsLoss1 == "died", 1, 0)

# Calculate the mean of ManualPCV for the specific CalfID
mean_manualPCV <- mean(ideal_data$ManualPCV[ideal_data$CalfID == "CA031210334"], na.rm = TRUE)
ideal_data$ManualPCV[ideal_data$CalfID == "CA031210334" & is.na(ideal_data$ManualPCV)] <- mean_manualPCV


# Fill missing Q.Strongyle.eggs based on CalfID group
ideal_data <- ideal_data %>%
  group_by(CalfID) %>%
  mutate(Q.Strongyle.eggs = ifelse(is.na(Q.Strongyle.eggs), mean(Q.Strongyle.eggs, na.rm = TRUE), Q.Strongyle.eggs)) %>%
  ungroup()

# Scale numerical variables
#num_vars <- c("Weight", "Age", "ManualPCV", "RecruitWeight", "Q.Strongyle.eggs")
#ideal_data[num_vars] <- scale(ideal_data[num_vars])

# Remove any rows with missing data
ideal_data <- ideal_data %>% na.omit()


# Convert categorical variables to factors
ideal_data <- ideal_data %>%
  mutate(
    CalfID = as.factor(CalfID),
    CalfSex = as.factor(CalfSex),
    Education = as.factor(Education),
    Distance_water = as.factor(Distance_water),
    ELISA_mutans = as.factor(ELISA_mutans),
    ELISA_parva = as.factor(ELISA_parva),
    Theileria.spp. = as.factor(Theileria.spp.),
    sublocation = as.factor(sublocation)
  )

summary(ideal_data)


# EDA: Visualize distribution of numerical variables
num_vars <- c("ADWG", "Weight", "ManualPCV", "Q.Strongyle.eggs", "RecruitWeight")
par(mfrow = c(2, 2))
for (var in num_vars) {
  ggplot(ideal_data, aes_string(x = var)) +
    geom_histogram(bins = 30, fill = "blue", alpha = 0.7) +
    labs(title = paste("Distribution of", var), x = var, y = "Frequency") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
}

# EDA: Check for missing values
missing_values <- colSums(is.na(ideal_data))
print(missing_values)

# EDA: Visualize relationships between variables
ggplot(ideal_data, aes(x = Weight, y = ADWG, color = CalfSex)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  labs(title = "Relationship between Weight and ADWG by Calf Sex", x = "Weight", y = "ADWG") +
  theme_minimal()

# EDA: Boxplot for categorical variable analysis
ggplot(ideal_data, aes(x = CalfSex, y = ADWG)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "ADWG by Calf Sex", x = "Calf Sex", y = "ADWG") +
  theme_minimal()

# Step 2: Fit the Mixed-Effects Model
mixed_model <- lmer(ADWG ~ CalfSex + Education + Distance_water +  RecruitWeight +
                      ManualPCV + Theileria.spp. + ELISA_parva + 
                      Q.Strongyle.eggs + (1 | CalfID), 
                    data = ideal_data, 
                    control = lmerControl(optimizer = "bobyqa")
                    )


# Step 3: Model Summary
mixed_model_summary <- summary(mixed_model)
print(mixed_model_summary)

# Step 4: Model Diagnostics
# Residuals vs Fitted plot
par(mfrow = c(2, 2))
plot(mixed_model)

# Check for multicollinearity using VIF
vif_values <- vif(mixed_model)
print(vif_values)

# Step 5: Predictions and Evaluation
# Store residuals in the data frame
ideal_data$residuals <- residuals(mixed_model)

# Calculate RMSE
rmse <- sqrt(mean(ideal_data$residuals^2))
cat("RMSE:", rmse, "\n")

# Calculate R-squared for mixed-effects models
r_squared <- r.squaredGLMM(mixed_model)
cat("R-squared (marginal):", r_squared[1], "\n")
cat("R-squared (conditional):", r_squared[2], "\n")


```


### LinearRegression
Removing Weight as a variable greatly increased the model's R^2, reduced the RMSE.
Introducing the optimizer made the model converge.

